### 2021-07-28 paper/localpsf.tex
edits to impulse response regular grid discussion

We compute impulse responses of $\mathcal{A}$ in batches by applying $\mathcal{A}$ to a small number Dirac combs. The batches of impulse responses, $\eta^b$, will ultimately be interpolated with weighting functions described in Section \ref{sec:weighting_functions} to form the desired approximation $\widetilde{\mathcal{A}}\approx \mathcal{A}$.  The Dirac comb, $\xi_b$, associated with a batch of sample points $S_b$ is the sum of Dirac distributions (point sources) centered at the points $x_i \in S_b$. That is,
\begin{equation*}
	\xi^b := \sum_{x_i \in S_b} \delta_{x_i}.
\end{equation*}
For each batch $S_b$, we compute as the action of $\mathcal{A}$ on the associated Dirac comb:
\begin{equation}
\label{eq:dirac_comb_H_action}
	\eta^b := \langle\mathcal{A}, \xi^b\rangle^*.
\end{equation}
By linearity, $\eta^b$ may be written as
\begin{equation}
\label{eq:phi_b}
	\eta^b = \left\langle\mathcal{A},\sum_{x_i \in S_b} \delta_{x_i}\right\rangle^* = \sum_{x_i \in S_b} \langle\mathcal{A}, \delta_{x_i}\rangle^* = \sum_{x_i \in S_b} \phi_x.
\end{equation}
Since the points $x_i$ are chosen so that the support of $\phi_i$ and the support of $\phi_j$ do not overlap (or do not overlap much) when $i \neq j$, from \eqref{eq:convolution_kernel} and \eqref{eq:phi_b} we have
\begin{equation}
\label{eq:varphi_eval}
	\varphi_{x_i}(z) = \phi_i(z+x_i) = \begin{cases}
		\eta^b(z+x_i), & z+x_i \in E_{x_i} \\
		0, & \text{otherwise}.
		\end{cases}
\end{equation}
for all $x_i \in S_b$. By performing one matrix-vector product, \eqref{eq:dirac_comb_H_action}, we recover $\varphi_{x_i}$ for every point $x_i \in S_b$. To minimize memory usage, the functions $\varphi_i$ are not stored individually. Instead, we store the functions $\eta^b$, and use formula \eqref{eq:varphi_eval} to perform function evaluations $z \mapsto \varphi_i(z)$ as needed.

@@

\subsection{Evaluating kernel entries of the approximation}
\label{sec:eval_matrix_entries}

We now describe how to efficiently evaluate $\widetilde{\Phi}(y,x)$ at an arbitrary point $(x,y) \in \Omega \times \Omega$. From \eqref{eq:kernel_entries}, we have
\begin{equation}
\label{eq:eval_y_x_000}
\AkerPc(y,x) = \sum_{i=1}^r w_i(x) \varphi_{x_i}(y-x),
\end{equation}
and from \eqref{eq:varphi_eval}, we have
\begin{equation}
\label{eq:kernel_entry_formua}
\varphi_{x_i}(y-x) = \begin{cases}
\eta^{b(i)}(y - x + x_i), & y - x + x_i \in E_{x_i} \\
0, & \text{otherwise},
\end{cases}
\end{equation}
where $b(i)$ denotes the batch index such that $x_i \in S^{b(i)}$. To compute $z=\AkerPc(y,x)$, we start with $z=0$, then loop through all batches $b$, and all sample points in each batch. For each sample point, we add one term in the sum in \eqref{eq:eval_y_x} to $z$, using \eqref{eq:kernel_entry_formua} as needed to evaluate $\varphi_{x_i}$. This is shown in Algorithm \ref{alg:eval_y_x}.

\begin{algorithm2e}
	\SetAlgoNoLine
	\SetKwInOut{Input}{Input}
	\SetKwInOut{Output}{Output}

	\Input{points $x \in \Omega$, $y \in \Omega$, \\sample point batches $S_1, \dots, S_\text{max\_batches}$, \\Dirac comb responses $\{\eta^b\}_{b=1}^{\text{max\_batches}}$, \\weighting functions $\{w_i\}_{i=1}^r$}
	\Output{Kernel entry $z = \widetilde{A}(y,x)$}

	$z \gets 0$

	\For{$b=1,2,\dots,\text{}max\_batches$}{
		\For{$x_i \in S_b$}{
			\If{$y - x + x_i \in E_{x_i}$}{
				$z \gets z + w_i(x)\eta^b(y-x+x_i)$
			}
		}
	}	
	\caption{Compute approximate kernel entry $z = \widetilde{A}(y,x)$}
	\label{alg:eval_y_x}
\end{algorithm2e}

>>>>

We compute impulse responses of $\mathcal{A}$ in batches by applying $\mathcal{A}$ to a small number Dirac combs. 
%The batches of impulse responses, $\eta^b$, will ultimately be interpolated with weighting functions described in Section \ref{sec:weighting_functions} to form the desired approximation $\widetilde{\mathcal{A}}\approx \mathcal{A}$.  
The Dirac comb, $\xi_b$, associated with a batch of sample points, $S_b$, is the sum of Dirac distributions (point sources) centered at the points $x_i \in S_b$. That is,
\begin{equation*}
	\xi^b := \sum_{x_i \in S_b} \delta_{x_i}.
\end{equation*}
For each batch $S_b$, we compute the action of $\mathcal{A}$ on the associated Dirac comb:
\begin{equation}
\label{eq:dirac_comb_H_action}
	\eta^b := \langle\mathcal{A}, \xi^b\rangle^*.
\end{equation}
By linearity, the resulting function $\eta^b$ may be written as
\begin{equation}
\label{eq:phi_b}
	\eta^b = \left\langle\mathcal{A},\sum_{x_i \in S_b} \delta_{x_i}\right\rangle^* = \sum_{x_i \in S_b} \langle\mathcal{A}, \delta_{x_i}\rangle^* = \sum_{x_i \in S_b} \phi_x.
\end{equation}
Since the points $x_i$ are chosen so that the support of $\phi_i$ and the support of $\phi_j$ do not overlap (or do not overlap much) when $i \neq j$, from \eqref{eq:convolution_kernel} and \eqref{eq:phi_b} we have
\begin{equation}
\label{eq:varphi_eval}
	\varphi_{x_i}(z) = \phi_i(z+x_i) = \begin{cases}
		\eta^b(z+x_i), & z+x_i \in E_{x_i} \\
		0, & \text{otherwise}.
		\end{cases}
\end{equation}
for all $x_i \in S_b$. By performing one matrix-vector product, \eqref{eq:dirac_comb_H_action}, we recover $\varphi_{x_i}$ for every point $x_i \in S_b$.

Later we will need to perform function evaluations $z \mapsto \varphi_{x_i}(z)$ for a large number of points $z \in \Omega$. Using \eqref{eq:varphi_eval} to perform this task requires evaluating the function $\eta_b$ at a large number of points $z+x_i$. In computations the functions $\eta^b$ will be finite element functions formed on an irregular mesh, and the points $z+x_i$ will typically not be gridpoints of the mesh. Evaluating a finite element function at an arbitrary point requires finding which cell in the mesh that the point is in, and then evaluating several polynomial functions to evaluate the finite element function at the point in that cell. While this procedure is asymptotically scalable, it tends to be slow in practice \cite{FEMEVALSLOW}. To speed up evaluation of the function $\varphi_{x_i}$, we interpolate $\varphi_{x_i}$ onto a regular rectilinear grid on a coordinate axis aligned box that just barely contains the ellipsoid $E_{x_i}$. The function $\varphi_{x_i}$ is transferred from the irregular mesh to the regular rectilinear grid by evaluating $\varphi_{x_i}$ at all gridpoints in the regular rectilinear grid using \eqref{eq:varphi_eval} and the slow finite element evaluation process described previously. Once $\varphi_{x_i}$ has been transferred to the regular rectilinear grid, further evaluations of $\varphi_{x_i}$ are performed extremely quickly using standard algorithms for regular rectilinear grid interpolation. To minimize the error associated with transferring the function $\varphi_{x_i}$ from the irregular mesh to the regular rectilinear grid, the density of the regular rectilinear grid should be denser than the local density of the finite element mesh. In our numerical results, the distance between neighboring points in the regular rectilinear grid is chosen to be half the minimum distance between any pair of finite element Lagrange nodes that are contained in the ellipsoid $E_{x_i}$. This procedure is illustrated in Figure FIG.

%To minimize memory usage, the functions $\varphi_i$ are not stored individually. Instead, we store the functions $\eta^b$, and use formula \eqref{eq:varphi_eval} to perform function evaluations $z \mapsto \varphi_i(z)$ as needed.

@@

%\subsection{Evaluating kernel entries of the approximation}
%\label{sec:eval_matrix_entries}
%
%We now describe how to efficiently evaluate $\widetilde{\Phi}(y,x)$ at an arbitrary point $(x,y) \in \Omega \times \Omega$. From \eqref{eq:kernel_entries}, we have
%\begin{equation}
%\label{eq:eval_y_x_000}
%\AkerPc(y,x) = \sum_{i=1}^r w_i(x) \varphi_{x_i}(y-x),
%\end{equation}
%and from \eqref{eq:varphi_eval}, we have
%\begin{equation}
%\label{eq:kernel_entry_formua}
%\varphi_{x_i}(y-x) = \begin{cases}
%\eta^{b(i)}(y - x + x_i), & y - x + x_i \in E_{x_i} \\
%0, & \text{otherwise},
%\end{cases}
%\end{equation}
%where $b(i)$ denotes the batch index such that $x_i \in S^{b(i)}$. To compute $z=\AkerPc(y,x)$, we start with $z=0$, then loop through all batches $b$, and all sample points in each batch. For each sample point, we add one term in the sum in \eqref{eq:eval_y_x} to $z$, using \eqref{eq:kernel_entry_formua} as needed to evaluate $\varphi_{x_i}$. This is shown in Algorithm \ref{alg:eval_y_x}.
%
%
%\begin{algorithm2e}
%	\SetAlgoNoLine
%	\SetKwInOut{Input}{Input}
%	\SetKwInOut{Output}{Output}
%	
%	\Input{points $x \in \Omega$, $y \in \Omega$, \\sample point batches $S_1, \dots, S_\text{max\_batches}$, \\Dirac comb responses $\{\eta^b\}_{b=1}^{\text{max\_batches}}$, \\weighting functions $\{w_i\}_{i=1}^r$}
%	\Output{Kernel entry $z = \widetilde{A}(y,x)$}
%
%	$z \gets 0$
%	
%	\For{$b=1,2,\dots,\text{}max\_batches$}{
%		\For{$x_i \in S_b$}{
%			\If{$y - x + x_i \in E_{x_i}$}{
%				$z \gets z + w_i(x)\eta^b(y-x+x_i)$
%			}
%		}
%	}	
%	\caption{Compute approximate kernel entry $z = \widetilde{A}(y,x)$}
%	\label{alg:eval_y_x}
%\end{algorithm2e}
